Abb Analysis

### 2017-02-06 abbDataAnalysis_Script.Rmd
re-working global imputation

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)

@@

  str.data$str.revenue <- str.data$bookings * str.data$med.rate

@@

  str.data$str.full.revenue <- (str.data$bookings * str.data$extr.par *
                                  str.data$med.rate)

@@

>>>>

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)
  library(kernlab)

@@

  str.data$str.act.revenue <- str.data$bookings * str.data$med.rate

@@

  str.data$str.ext.revenue <- (str.data$bookings * str.data$extr.par *
                                  str.data$med.rate)

  str.data$occ.rate <- 1 - (str.data$block.rate + str.data$avail.rate)

  str.data$pot.occ.rate <- str.data$bookings / (366 - (366 * str.data$block.rate))

  str.data$str.pot.revenue <- (str.data$pot.occ.rate * 366 * str.data$med.rate)

@@

### Host types

  ggplot(str.data, aes(x=block.rate)) + 
  geom_density(fill=str.col, color=str.col) +
  ylab('Frequency\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0,1,.2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Host Block Rate Frequency\n') +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

  ggplot(str.data, aes(x=block.rate, y=nbr.block*extr.par)) + 
  geom_point(fill=str.col, color=str.col, size=.2) +
  ylab('Frequency\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Host Block Rate Frequency\n') +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

str.data$blockpertime <- round(str.data$nbr.block * str.data$extr.par, 1)

str.data$host.type <- 'Unknown'
str.data$host.type[str.data$blockpertime > 12] <- 'Multi-Platform User'
str.data$host.type[str.data$block.rate <= .25] <- 'Profit Seeker'
str.data$host.type[str.data$block.rate >= .75] <- 'Opportunistic Sharer'
str.data$host.type <- factor(str.data$host.type,
                             levels=c('Profit Seeker', 'Opportunistic Sharer', 
                                      'Multi-Platform User', 'Unknown'))

ggplot(str.data, 
       aes(x=block.rate, y=blockpertime, color=host.type)) + 
  geom_point(alpha=.5, size=.8, shape=15) +
  scale_color_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))

ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none')

## Global Market Analysis

We now analyze the full metro market to test which properties are more profitable as short-term rentals.

We start by specifying the imputation models for the LTR and the Airbnb datasets

  ltr.mod.spec <- formula(log(event.price) ~ as.factor(type) + 
                            as.factor(bedbath) + as.factor(suburb) + 
                            as.factor(ltr.month))

We then employ a custom function to cross impute rates and rents for the two datasets.  The 'clip.field' is a spatial fixed effects that add spatial recognition to the models.

  imp.rents <- imputeRents(ltr.df=ltr.data, 
                           str.df=str.data, 
                           mod.spec=ltr.mod.spec,
                           clip.field='suburb')

  str.glob <- merge(str.data, imp.rents$str, by='property.id')

Next we cross impute days on market and occupancy rates.

  # Apply dom to abb
  str.glob$imp.dom <- imputeDOM(str.glob, ltr.data, calc.type='median')

HERE HERE HERE

With the cross-imputed DOMs and occupancy rates we now make the cross-tenure imputations of revenues and add them back to the original datasets.

  impx.revs <- revenueEngine(str.df=ltr.imp, 
                             ltr.df=str.imp,
                             rate.field='imp.rate',
                             rent.field='imp.rent',
                             occ.field='imp.occ',
                             dom.field='imp.dom')

  str.imp$ltr.imp.revenue <- impx.revs$ltr
  ltr.imp$str.imp.revenue <- impx.revs$str

Next, we compare the revenues from the two and extract the new datasets.

  comp.revs <- compareRevenues(str.imp)

  str.imp <- merge(str.imp, comp.revs, by='property.id')

### Analyze Results

First we make a comparison table of the results, broken down by submarket.

  mrkt.table <- createCompTable(str.imp, 'sub.mrkt')

  mrkt.table$ID <- factor(mrkt.table$ID, 
                          levels=c('city-core', 'city', 'suburban', 'rural', 'beach'))

We plot these results in a 2x2 matrix to show the preference likelihoods under the 4 possible scenarios.

  sm.col <- abb.col[c(1, 3, 6, 5, 2)]

  twotwo.bar.plot <- 
    ggplot(mrkt.table, aes(x=ID, weights=Var, fill=ID)) + 
    geom_bar() +
    facet_grid(~rev.type) +
    scale_fill_manual(values=sm.col) +
    xlab('') +
    ylab('% of properties where Airbnb is more profitable') +
    scale_y_continuous(breaks=c(0,.25,.5,.75,1),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none')

Next, we plot the \% of properties that have short-term preference versus the occupancy rate

  rawocc.pplot <- makePrefPlot(str.imp,
                               x.field='pot.occ.rate',
                               y.field='str.ext.pref',
                               group.field='host.type',
                               smooth=TRUE,
                               smooth.span=.75)
  rawocc.pplot <- rawocc.pplot +
    xlab('\nOccupancy Rate') +
    ylab('\n% of Properties where Airbnb is more profitable') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_color_manual(values=sm.col)

Looking at the distribution of occupancy rates, we that it is not uniform in any of the submarkets.

  str.revs$sub.mrkt <- factor(str.imp$sub.mrkt, 
                         levels=c('city-core', 'city', 'suburban', 'rural', 'beach'))
  ggplot(str.imp, aes(x=pot.occ.rate, group=sub.mrkt, color=sub.mrkt, fill=sub.mrkt)) +
    geom_density(alpha=.3) +
    facet_wrap(~sub.mrkt) +
    scale_color_manual(values=sm.col) +
    scale_fill_manual(values=sm.col) +
    scale_x_continuous(breaks=seq(0, 1, by=.5),
                       labels=c('0%', '50%', '100%')) +
    ylab('') +
    xlab('Occupancy Rate') +
    theme(legend.position='none',
          axis.line=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.y=element_blank())

We then calculate quantile values for rates and replot the charts

  str.imp$occ.qtl <- makeWtdQtl(str.imp$occ.rate, 
                                      return.type='rank') 
  str.imp$rate.qtl <- makeWtdQtl(str.imp$med.rate, 
                                  return.type='rank')

  ## Make quartile location plot

  qtlocc.pplot <- makePrefPlot(str.imp,
                               x.field='occ.qtl',
                               y.field='str.act.pref',
                               group.field='sub.mrkt',
                               smooth=TRUE,
                               smooth.span=.75)
  qtlocc.pplot <- qtlocc.pplot +
    xlab('\nQualtile of Occupancy Rate') +
    ylab('\n% of Properties where Airbnb is more profitable') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_color_manual(values=sm.col)

Next we build a heat map that shows which combinations of occ rate and nightly rate are most likely to result in short-term preference.

  rate.hm <- makeHeatMap(str.imp,
                x.field='occ.rate',
                y.field='nightly.rate',
                color.field='str.act',
                bins=c(.05, 25),
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))

  rate.hm <- rate.hm +
    xlab('\n Occupancy Rate') +
    ylab('\n Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom')

To make the distinction more clear, we employ a Support Vector Machine to classify the areas into short and long term preference.

  rate.hm.svm <- makeHeatMap(str.revs,
                         x.field='occ.rate',
                         y.field='nightly.rate',
                         color.field='str.act',
                         bins=c(.05, 25),
                         svm=T, 
                         alpha.count=F,
                         add.points=T,
                         fill.colors=c(abb.col[1], abb.col[5]))

  rate.hm.svm <- rate.hm.svm +
    xlab('\n Occupancy Rate') +
    ylab('\n Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom')

Again, due to the non-uniformity of distribution of these two axis, we convert to quantiles and re-plot.

  qtl.hm <- makeHeatMap(str.revs,
                         x.field='occ.qtl',
                         y.field='rate.qtl',
                         color.field='str.act',
                         bins=c(5, 5),
                         svm=F, 
                         alpha.count=T,
                         add.points=T,
                         fill.colors=c(abb.col[1], abb.col[5]))

  qtl.hm <- qtl.hm +
    xlab('\n Quantile of Occupancy Rate') +
    ylab('\n Quantile of Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    theme(legend.position='bottom')

  qtl.hm.svm <- makeHeatMap(str.revs,
                        x.field='occ.qtl',
                        y.field='rate.qtl',
                        color.field='str.act',
                        bins=c(5, 5),
                        svm=T, 
                        alpha.count=F,
                        add.points=T,
                        fill.colors=c(abb.col[1], abb.col[5]))

  qtl.hm.svm <- qtl.hm.svm +
    xlab('\n Quantile of Occupancy Rate') +
    ylab('\n Quantile of Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    theme(legend.position='bottom')

Finally, we calculate the areas of the heat maps and combine into a table.

  svm.rate <- makeSVM(str.revs,
                     x.field='occ.rate',
                     y.field='nightly.rate',
                     z.field='str.act',
                     svm.type='C-svc',
                     svm.kernel='polydot',
                     poly.degree=4,
                     expand.factor=100)

  svm.qtl <- makeSVM(str.revs,
                      x.field='occ.qtl',
                      y.field='rate.qtl',
                      z.field='str.act',
                      svm.type='C-svc',
                      svm.kernel='polydot',
                      poly.degree=4,
                      expand.factor=100)

  market.ratio <- data.frame(type=c('rate', 'qtl'),
                             actual=rep(mean(str.revs$str.act), 2),
                             fitted=c(mean(svm.rate$orig$fitted),
                                      mean(svm.qtl$orig$fitted)),
                             svm=c(mean(svm.rate$pred$pred),
                                   mean(svm.qtl$pred$pred)))
```

### Submarkets

We now take the analytical process just completed on the entire metro and apply it to various submarkets.  We use a custom function that combines all the processes above into a single command.

#### Property Type

We start by comparing apartments to houses.  First we set a new imputation model specification

  # Set model specifications
  ltr.mod.spec <- formula(log(event.price) ~ as.factor(bedbath) + as.factor(suburb) + 
                            as.factor(ltr.month))
  str.mod.spec <- formula(log(nightly.rate) ~ as.factor(bedbath) + as.factor(suburb))

We then recompute the market analysis for each property type.

  # Select data
  str.apt <- which(str.revs$type == 'Apartment')
  ltr.apt <- which(ltr.revs$type == 'Apartment')

  # Estimate House Results  
  house <- fullMarketAnalysis(ltr.df=ltr.revs[-ltr.apt,],
                              str.df=str.revs[-str.apt, ],
                              ltr.mod.spec=ltr.mod.spec,
                              str.mod.spec=str.mod.spec,
                              clip.field='suburb',
                              market.field='none',
                              mrkt.col=sm.col,
                              heat.col=c(abb.col[1], abb.col[5]))

   apt <- fullMarketAnalysis(ltr.df=ltr.revs[ltr.apt,],
                             str.df=str.revs[str.apt, ],
                             ltr.mod.spec=ltr.mod.spec,
                             str.mod.spec=str.mod.spec,
                             clip.field='suburb',
                             market.field='sub.mrkt',
                             mrkt.col=sm.col,
                             heat.col=c(abb.col[1], abb.col[5]))
